--
title: "Adaptive Sampling simulations"
author: "Corey Williams"
date: "`r Sys.Date()`"
output: html_document
---

# Motivation for adaptive sampling
In many scenarios we would like to be able to adaptively increase sampling effort when certain observed values are of interest. This would be especially beneficial in the case where we are try to observe rare events that are likely to be clustered together. For example a rare plant species that has very particular growing conditions. This would be the case where it may be very rare to see the plant but if we see one we see 100 and it is desired to be able to sample all of the plants in the cluster.

# Simulations
First I'll write a few functions that I think will be helpful in making this adaptive sampling demonstration. I want to recreate the example in Adaptive cluster Sampling (1990) where there are points distributed in clusters in an area. The area is split into a grid, then grid cells are chosen as the primary sampling unit.

Goals:

- choose a number of clusters
  - easy enough, this can just be rpois(1,n), this will choose a number of clusters randomly based on some mean. We could use any discrete distribution to decide this
- generate points in those clusters
  - decide how many points are in each cluster, this it is probably actually most appropriate to use a poisson distribution since it will be a count.
 - at this step I also need to choose the locations of the centers, runif(ncenters,0,20)
  - I can then generate points normally distributed around these centers using rnorm(ncenters*2,mu,sd)
- choose a sample of initial grid cells
  - sample(1:ncells,n)
- expand if the cell next to it is occupied
  - need a list of currently occupied cells and a way to determine adjacent cells. maybe just a vector called occupied
  - function takes vector of neighbours, checks against vector of occupied, returns neighbors in occupied


## write helpful functions

```{r}
require(tidyverse)
```

### make clusters
```{r}
make_clusters<-function(grid_size=20, nclusters=3, avg_size=20){
  # create a list of clusters.
  centers<-split(runif(nclusters*2,0,grid_size),seq(nclusters))
  # get the three cluster sizes
  sizes<-rpois(3,avg_size)
  # get list of matrices of the centers
  center_dfs<-map2(sizes,centers,
                   ~ kronecker(matrix(rep(1,.x),ncol=1), matrix(.y,ncol=2)))
                              # I actually used the kronecker product holy crap!
  # get coordinates of locations
  locations<-sizes %>% # use sizes
    map(~data.frame(matrix(rnorm(.x*2),ncol=2))) %>% # make list of df of changes from centers
    map2(.y=center_dfs,~.x+.y) %>% # add the changes on to the center
    map2(.y=seq(nclusters), ~ mutate(.x, Group=paste("Cluster",.y)))%>%
    bind_rows()
}
```

### make clusters
```{r}
plot_clusters<-function(cluster_df, grid_size=20,samp=NULL){
  # create a plot of the clusters
  p<-ggplot(cluster_df, aes(x=X1, y=X2, color=Group))+
    geom_point()+
    scale_x_continuous(breaks=seq(grid_size))+
    scale_y_continuous(breaks=seq(grid_size))+
    coord_cartesian(xlim=c(0,grid_size), ylim=c(0,grid_size))+
    coord_equal()
  if(!is.null(samp)){
    # this will highlight tiles that are sampled
    # samp is the dataframe containing the coordinates for tiles
    p<-p+geom_rect(data=samp, 
                   aes(x=X, y=Y, xmin=X-1, xmax=X, ymin=Y-1, ymax=Y),
                   color="black",
                   fill=NA)
  }
  p
}
```

### Check whether a tile is occupied
```{r}
is_occupied<-function(tile=c(1,1),df){
  # check if a tile is occupied tile location is given as c(row,column)
  # give a dataframe with coordinates of point
  # check if there are any points in xrange & yrange at the same time
  #sum((df[,1]<tile[2] & df[,1]>tile[2]-1) * (df[,1]<tile[2] & df[,1]>tile[2]-1))>0
  
  sum((df[,1]<tile[1]&df[,1]>tile[1]-1) & (df[,2]<tile[2] & df[,2]>tile[2]-1))>0
}
```


### get the neighbours of a tile
```{r}
get_neighbours<- function(tile=c(1,1), hard_border=TRUE,grid_size=20){
  # gets a list of neighbouring tiles
  neighbours<-list(c(tile[1]-1,tile[2]),
                   c(tile[1],tile[2]-1),
                   c(tile[1]+1,tile[2]),
                   c(tile[1],tile[2]+1))
  if(hard_border){
    # do we want to include neighbours outside of the grid?
    # returns the neighbours that are only within the border
    neighbours<-neighbours[map_lgl(neighbours, ~prod(c(.x>0,.x<=grid_size)))]
  }
  # returns list of neighbours for a given tile 
  neighbours
}
```

### choose tiles for sample

```{r}
get_tiles<-function(grid_size=20,n1=10,...){
  # this function returns a sample of n1 tiles using a square grid_size grid
  # get a tile
  tiles<-sample(1:grid_size^2,n1,...)
  # convert tile numbers into X and Y
  samp<-data.frame(X=tiles%%grid_size, Y=(tiles-1)%/%grid_size+1)
  samp$X[samp$X==0]<-grid_size
  samp
}

```

### perform the simulation
```{r}
simulate_one<-function(nclusters=3, grid_size=20,... ){
  # generate clusters
  points<-make_clusters()
  # choose the starting grid cells
  sample_tiles<-get_tiles(...)
  # save a copy to check against for updates
  temp<-sample_tiles
  # check whether or not they are occupied based on the clusters
  occupied<-apply(sample_tiles,1,function(x) is_occupied(x,points))
  # find the neighbours of the occupied points
  neighbours<-apply(sample_tiles[occupied,], 1, get_neighbours)%>% # get neighbours
    bind_rows() %>% # turn list of neighbours into tibble
    as.data.frame() # into dataframe
  # update sample tiles to include neighbours
  sample_tiles<-rbind(sample_tiles,neighbours) %>%
    unique()
  
  # keep looping until sample_tiles does not grow
  while(dim(temp)[1]!=dim(sample_tiles)[1]){ 
    # save a copy to check against for updates
    temp<-sample_tiles
    # check whether or not they are occupied based on the clusters
    occupied<-apply(sample_tiles,1,function(x) is_occupied(x,points))
    # find the neighbours of the occupied points
    neighbours<-apply(sample_tiles[occupied,], 1, get_neighbours)%>% # get neighbours
      bind_rows() %>% # turn list of neighbours into tibble
      as.data.frame() # into dataframe
    # update sample tiles to include neighbours
    sample_tiles<-sample_tiles %>%
      bind_rows(neighbours) %>%
      unique()
  }
  return(list(sample_tiles=cbind(sample_tiles,occupied),points=points))
}


```

## Demonstrating the function works to generate a sample
```{r}
sample_one<-simulate_one(n1=10,nclusters=3, grid_size=20)
plot_clusters(sample_one$points,samp=sample_one$sample_tiles)
```


When trying to write this simulation the largest issue I ran in to was determining how to treat points that lie outside of the grid. Should they be included in the sample? should they be ignored? should I change the generating mechanism to force points to be bounded by the edges? There are lots of things you need to specify in regards to how the points are generated. This seems like it could also be an issue that comes up in practical situations as well, for example the case where you have a defined area you are allowed to collect samples from but the thing you are measuring can occur up to and outside of that area. It seems like we would be underestimating the average number in the greater population if those are excluded from the cluster since it means we are underestimating the cluster size.


# Finding the HT and HH Estimators

## Vocabulary
 - neighborhood: the collection of units that are immediately included in the sample if a given unit is included. This relationship is symmetric and is typically (but not necessarily) geographic.
 - Cluster: the collection of all the units that are observed under the design as a result of initial selection of unit $i$
 - Network: selection of any unit within the network would lead to the inclusion in the sample of every other unit in the network.
 - edge unit: any unit not satisfying the condition but in the neighborhood of one that does 

## Hansen-Hurwitz
 - Let $\Psi_k$ be the network that includes unit $k$ and $m)k$ be the number of units in that network. Any unit not satisfying the criterion is size $1$. 
 - Let $\bar y_k^*=(m_k)^{-1}\sum_{j\in\Psi_k}y_j$ represent the average number of observations in the network that includes the $k$th unit  of the initial sample
 - The modified Hansen-Hurwitz estimators given as $t_{HH^*}=n_1^{-1}\sum^{n_1}_{k=1}\bar y_k^*$

## Hovritz-Thompson


# Scratch work/debugging
```{r eval=FALSE}
grid_size<-20
nclusters<-3
n1<-10
# create clusters
points<-make_clusters()
# plot them
plot_clusters(points)
# choose the starting grid cells
sample_tiles<-matrix(sample(1:grid_size,n1*2),ncol=2)%>%
  as.data.frame()

plot_clusters(points,samp=sample_tiles)

cbind(sample_tiles,apply(sample_tiles,1,is_occupied, df=points))


```


```{r}
# Example dataframe
df <- data.frame(
  A = c(1, 2, 3),
  B = c(4, 5, 6),
  C = c(7, 8, 9)
)

# Example vector to add
addition_vector <- c(10, 20, 30)

# Add the vector to every row of the dataframe
result <- df + addition_vector

# View the result
print(result)

```




```{r}
getwd()
```









